---
title: "cowplot package"
subtitle: "Gracia Larsen-Schmidt"
format:
  revealjs: 
    slide-number: true
    chalkboard: 
      buttons: false
    logo: cowplot_graphic.png
editor_options:
  chunk_output_type: console
---

```{r}
#loading in libraries
#install.packages(cowplot)
library(cowplot)
library(knitr)
library(tidyverse)
```

```{r}
#bringing in dataset
outer_space_objects <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-04-23/outer_space_objects.csv')
```

```{r include = FALSE}
# scatterplot of objects sent into space over time 
all_entities_over_time <- outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")
all_entities_over_time
```

```{r include = FALSE}
# Creating a table of overall totals to find out which entities sent out the most 1957-2023. "World" is the overall, combined total of objects launched into outer space. 
total_count <- outer_space_objects |>
  select(-c(Year, Code)) |>
  group_by(Entity) |>
  mutate(total = sum(num_objects)) |>
  select(total) |>
  unique() |>
  arrange(desc(total))
total_count

kable(head(total_count, 10))
```

```{r include = FALSE}
# This graph maps the trajectory of number of objects sent out over time using the non-"World" entities above that have sent out the most total since 1957 (US, Russia, China, UK, and Japan) 
top_5_over_time <- outer_space_objects |>
  filter(Entity %in% c("United States","Russia", "China", "United Kingdom", "Japan")) |>
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = fct_reorder2(Entity, Year, num_objects))) +
  geom_point() +
  labs(title = "Entities Launching the Most Objects Into Space Over Time", 
       y = "Number of Objects",
       color = "Entity")
top_5_over_time
```

```{r include = FALSE}
# Bar graph of total objects launched into space (top 5 entities)
bar_graph_totals <- total_count |>
  filter(Entity %in% c("United States","Russia", "China", "United Kingdom", "Japan")) |>
  ggplot(aes(x = fct_reorder(Entity, total, .desc = TRUE), y = total, fill = fct_reorder2(Entity, Entity, total))) +
  geom_bar(stat = "identity") +
  labs(title = "Total Objects Launched Into Space by Top 5 Object-Launching Entities", 
       x = "Entity", 
       y = "Total Objects Launched", 
       fill = "Entity")

bar_graph_totals
```

## Why cowplot?

  - Easy to use
  - Works well with ggplot
  - Useful for producing and organizing figures for projects and publications

::: footer
Learn more: [cowplot introductory vignette](https://wilkelab.org/cowplot/articles/introduction.html)
:::

## Data {.smaller}

::: columns
::: {.column width="45%"}
<br/> 

#### *Objects Launched into Space*
This data is from [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday). The dataset, titled "outer_space_objects," includes four variables: Entity, Code, Year (ranging from 1957 to 2023), and a variable indicating the number of objects launched into space in a given year by a given entity. 
:::

::: {.column width="3%"}
:::

::: {.column width="52%"}

```{r}
kable(head(outer_space_objects)[,c("Entity",	"Code", "Year", "num_objects")])
```
:::
:::

::: footer
Data Source: [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/5c99ca51eed16df5aaa81a814ac0dcd287a1a315/data/2024/2024-04-23/readme.md)
:::

## plot_grid() {.smaller .scrollable}

::: panel-tabset
### Unaligned

```{r}
outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")

outer_space_objects |>
  filter(Entity %in% c("United States","Russia", "China", "United Kingdom", "Japan")) |>
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = fct_reorder2(Entity, Year, num_objects))) +
  geom_point() +
  labs(title = "Entities Launching the Most Objects Into Space Over Time",
       y = "Number of Objects",
       color = "Entity")
```

### Aligned

```{r}
all_entities_over_time <- outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")

top_5_over_time <- outer_space_objects |>
  filter(Entity %in% c("United States","Russia", "China", "United Kingdom", "Japan")) |>
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = fct_reorder2(Entity, Year, num_objects))) +
  geom_point() +
  labs(title = "Entities Launching the Most Objects Into Space Over Time",
       y = "Number of Objects",
       color = "Entity")

plot_grid(all_entities_over_time, top_5_over_time, ncol = 1, align = "v")
```
:::

::: footer
Learn more: [Aligning by Axis](https://wilkelab.org/cowplot/articles/aligning_plots.html)
:::

## Code

```{.r code-line-numbers="1-17|17"}
all_entities_over_time <- outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")

top_5_over_time <- outer_space_objects |>
  filter(Entity %in% c("United States","Russia", "China", "United Kingdom", "Japan")) |>
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = fct_reorder2(Entity, Year, num_objects))) +
  geom_point() +
  labs(title = "Entities Launching the Most Objects Into Space Over Time",
       y = "Number of Objects",
       color = "Entity")

plot_grid(all_entities_over_time, top_5_over_time, ncol = 1, align = "v")
```

::: footer
Learn more: [Aligning by Axis](https://wilkelab.org/cowplot/articles/aligning_plots.html)
:::


## ggdraw function: draw_label() {.smaller}
```{r}
all_entities_over_time <- outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")

ggdraw(all_entities_over_time) +
  draw_label("Draft", 
             color = "#FF3333", 
             size = 75, 
             angle = 35, 
             alpha = .25, 
             x = .5, 
             y = .55)
```

::: footer
Learn more: [Drawing With and on Plots](https://wilkelab.org/cowplot/articles/drawing_with_on_plots.html)
:::

## Code

```{.r code-line-numbers="1-14|7|8-14"}
all_entities_over_time <- outer_space_objects |>
  ggplot(aes(x = Year, y = num_objects)) +
  geom_point() +
  labs(title = "Number of Objects Launched Into Space Over Time",
       y = "Number of Objects")

ggdraw(all_entities_over_time) +
  draw_label("Draft", 
             color = "#FF3333", 
             size = 75, 
             angle = 35, 
             alpha = .25, 
             x = .5, 
             y = .55)
```

::: footer
Learn more: [Drawing With and on Plots](https://wilkelab.org/cowplot/articles/drawing_with_on_plots.html)
:::

## ggdraw function: draw_image()
```{r}
#install.packages("magick")
library(magick)

ggdraw() + 
  draw_plot(top_5_over_time) +
  draw_image("cowplot_graphic.png", 
             x = 1, 
             y = 1, 
             hjust = 1, 
             vjust = 1, 
             halign = 1, 
             valign = 1, 
             width = .1)
```
::: footer
Learn more: [Drawing With and on Plots](https://wilkelab.org/cowplot/articles/drawing_with_on_plots.html)
:::

## Code
```{.r code-line-numbers="1-2|4|5|6-13"}
#install.packages("magick")
library(magick)

ggdraw() + 
  draw_plot(top_5_over_time) +
    draw_image("cowplot_graphic.png", 
               x = 1, 
               y = 1, 
               hjust = 1, 
               vjust = 1, 
               halign = 1, 
               valign = 1, 
               width = .1)
```

::: footer
Learn more: [Drawing With and on Plots](https://wilkelab.org/cowplot/articles/drawing_with_on_plots.html)
:::

## theme(), get_legend(), and plot_grid() {.smaller .scrollable}
::: panel-tabset

### Without cowplot
```{r}
bar_graph_totals <- total_count |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  ggplot(aes(x = fct_reorder(Entity, total, .desc = TRUE), 
             y = total, 
             fill = fct_reorder2(Entity, Entity, total))) +
  geom_bar(stat = "identity") +
  labs(title = "Total Objects Launched 1957-2013", 
       x = "Entity", 
       y = "Total Objects", 
       fill = "Entity")
bar_graph_totals


top_5_over_time_faceted <- outer_space_objects |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  mutate(Entity = factor(Entity, levels = c("United States",
                                            "Russia", 
                                            "China", 
                                            "United Kingdom", 
                                            "Japan"))) |> 
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = Entity)) +
  geom_point() +
  facet_wrap(~Entity, nrow = 1) +
  labs(title = "Entities Launching Over Time",
       y = "Number of Objects",
       color = "Entity")
top_5_over_time_faceted
```

### Using cowplot {.smaller}
```{r}
bar_graph_totals <- total_count |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  ggplot(aes(x = fct_reorder(Entity, total, .desc = TRUE), 
             y = total, 
             fill = fct_reorder2(Entity, Entity, total))) +
  geom_bar(stat = "identity") +
  labs(title = "Total Objects Launched 1957-2023",
       x = "", 
       y = "Total Objects", 
       fill = "Entity")


top_5_over_time_faceted <- outer_space_objects |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  mutate(Entity = factor(Entity, levels = c("United States", 
                                            "Russia", 
                                            "China", 
                                            "United Kingdom", 
                                            "Japan"))) |> 
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = Entity)) +
  geom_point() +
  facet_wrap(~Entity, nrow = 1) +
  labs(title = "Objects Launched Over Time",
       y = "Number of Objects")


pcol1 <- plot_grid(bar_graph_totals + theme(legend.position="none"),
                   top_5_over_time_faceted + theme(legend.position="none"),
                   labels = c('A', 'B'), 
                   label_size = 15, 
                   ncol = 1,
                   align = "v")

legend <- get_legend(bar_graph_totals + theme(legend.box.margin = margin(0, 0, 0, 10)))

plot_grid(pcol1, 
          legend, 
          nrow = 1,
          rel_widths = c(3, .5))

```

:::

::: footer
Learn more: [Themes](https://wilkelab.org/cowplot/articles/themes.html), [Shared Legends](https://wilkelab.org/cowplot/articles/shared_legends.html), and [Arranging Plots in a Grid](https://wilkelab.org/cowplot/articles/plot_grid.html)
:::

## Code {.smaller .scrollable}
```{.r code-line-numbers="1-49|37|38|39|40|41|42|44|46|47|48|49"}
bar_graph_totals <- total_count |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  ggplot(aes(x = fct_reorder(Entity, total, .desc = TRUE), 
             y = total, 
             fill = fct_reorder2(Entity, Entity, total))) +
  geom_bar(stat = "identity") +
  labs(title = "Total Objects Launched 1957-2023",
       x = "", 
       y = "Total Objects", 
       fill = "Entity")


top_5_over_time_faceted <- outer_space_objects |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  mutate(Entity = factor(Entity, levels = c("United States", 
                                            "Russia", 
                                            "China", 
                                            "United Kingdom", 
                                            "Japan"))) |> 
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = Entity)) +
  geom_point() +
  facet_wrap(~Entity, nrow = 1) +
  labs(title = "Objects Launched Over Time",
       y = "Number of Objects")


pcol1 <- plot_grid(bar_graph_totals + theme(legend.position="none"),
                   top_5_over_time_faceted + theme(legend.position="none"),
                   labels = c('A', 'B'), 
                   label_size = 15, 
                   ncol = 1,
                   align = "v")

legend <- get_legend(bar_graph_totals + theme(legend.box.margin = margin(0, 0, 0, 10)))

plot_grid(pcol1, 
          legend, 
          nrow = 1,
          rel_widths = c(3, .5))

```

::: footer
Learn more: [Themes](https://wilkelab.org/cowplot/articles/themes.html), [Shared Legends](https://wilkelab.org/cowplot/articles/shared_legends.html), and [Arranging Plots in a Grid](https://wilkelab.org/cowplot/articles/plot_grid.html)
:::

## {.smaller}
```{r}
bar_graph_totals <- total_count |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  ggplot(aes(x = fct_reorder(Entity, total, .desc = TRUE), 
             y = total, 
             fill = fct_reorder2(Entity, Entity, total))) +
  geom_bar(stat = "identity") +
  labs(title = "Total Objects Launched 1957-2023",
       x = "", 
       y = "Total Objects", 
       fill = "Entity")


top_5_over_time_faceted <- outer_space_objects |>
  filter(Entity %in% c("United States",
                       "Russia", 
                       "China", 
                       "United Kingdom", 
                       "Japan")) |>
  mutate(Entity = factor(Entity, levels = c("United States", 
                                            "Russia", 
                                            "China", 
                                            "United Kingdom", 
                                            "Japan"))) |> 
  group_by(Entity) |>
  mutate(Entity = Entity) |>
  ggplot(aes(x = Year, y = num_objects, color = Entity)) +
  geom_point() +
  facet_wrap(~Entity, nrow = 1) +
  labs(title = "Objects Launched Over Time",
       y = "Number of Objects")


pcol1 <- plot_grid(bar_graph_totals + theme(legend.position="none"),
                   top_5_over_time_faceted + theme(legend.position="none"),
                   labels = c('A', 'B'), 
                   label_size = 15, 
                   ncol = 1,
                   align = "v")

legend <- get_legend(bar_graph_totals + theme(legend.box.margin = margin(0, 0, 0, 10)))

plot_grid(pcol1, 
          legend, 
          nrow = 1,
          rel_widths = c(3, .5))

```

::: footer
Learn more: [Themes](https://wilkelab.org/cowplot/articles/themes.html), [Shared Legends](https://wilkelab.org/cowplot/articles/shared_legends.html), and [Arranging Plots in a Grid](https://wilkelab.org/cowplot/articles/plot_grid.html)
:::

## Questions?