---
title: "Mini Project 1"
format:
  pdf: default
editor_options: 
  chunk_output_type: console
---

https://data.hrsa.gov/maps/mchb

```{r}
library(tidyverse)
library(dplyr)
```

```{r}
cesarean_delivery <- read_csv("~/Desktop/Cesarean Delivery - All (%) 2020-2022 - Output for External Tool v2/Cesarean Delivery - All (%) 202-Table 1.csv") 

cesarean_delivery
```


```{r}
#loading in info to draw US states for geom_polygon
library(maps)
states_polygon <- as_tibble(map_data("state")) |>
  select(region, group, order, lat, long)

# See what the state (region) levels look like in states_polygon
unique(states_polygon$region)

#loading in info to draw US states for geom_sf and leaflet
library(sf)
states_sf <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  select(name, geometry)

# See what the state (name) levels look like in states_sf
unique(states_sf$name)

c_delivery <- as_tibble(cesarean_delivery) |>
  filter()

# See what the state (State) levels look like in c_delivery
unique(c_delivery$State)
```


```{r}
# Matching the key values
library(lubridate)

states_polygon <- states_polygon |>
  mutate(region = str_to_title(region))
states_polygon
```

```{r}
# Checking what states/regions need to be filtered out
c_delivery |>
  anti_join(states_polygon, by = c("State" = "region")) |>
  count(State)

c_delivery |> 
  anti_join(states_sf, by = c("State" = "name")) |>
  count(State)
```

```{r}
# Filtering 
c_delivery <- c_delivery |>
  filter(!(State %in% c("alaska",
                       "american samoa",
                       "federated states of micronesia",
                       "guam",
                       "hawaii",
                       "marshall islands",
                       "northern mariana islands",
                       "puerto rico",
                       "republic of palau",
                       "u.s minor islands",
                       "u.s. virgin islands")))
c_delivery
```


```{r}
# Merging for static
c_del_polygon <- c_delivery |>
   group_by(State) |> 
  summarise(mean_c_section_delivery = round(mean(`Cesarean Delivery - All (%) 2020-2022`), digits = 1), 
            mean_urban_rural = round(mean(`NCHS Urban-Rural Code`, na.rm = TRUE), digits = 0)) |>
  right_join(states_polygon, by = c("State" = "region"))
c_del_polygon
```


```{r}
#Merging for interactive
temp <- c_delivery |>
   group_by(State) |>
  summarise(mean_c_section_delivery = round(mean(`Cesarean Delivery - All (%) 2020-2022`), digits = 1),
            mean_urban_rural = round(mean(`NCHS Urban-Rural Code`, na.rm = TRUE), digits = 0))

c_del_sf_1 <- states_sf |>
  left_join(temp, by = c("name" = "State"))
```


# Average Percent Cesarian Deliveries By State 2020-2022 (static plot)
```{r}
library(viridis)
c_del_polygon |>
  ggplot(mapping = aes(x = long, y = lat, group = group)) + 
    geom_polygon(aes(fill = mean_c_section_delivery), color = "black") + 
    labs(fill = "Average % Cesarian Deliveries", caption = "Data Source: https://data.hrsa.gov/maps/mchb", title = "Average % Cesarian Deliveries By State (2020-2022)") +
    coord_map() + 
   theme_void() +  
   scale_fill_viridis() 
```
- This graph is a map of the continental United States color coded with state-level data of the average percent of deliveries which were cesarean (c-sections) in the years 2020-2022, aggregated from county-level data. The average percent of deliveries which were cesarean varies from approximately 24% to 38%. The graph highlights that the South tends to have higher rates of cesarian deliveries, especially Louisiana and Mississippi, while the West and upper-western Midwest tend to have lower rates of cesarian sections. 


# Average County Urban-Rural Code by State (static plot)
```{r}
table(c_del_polygon$mean_urban_rural)

c_del_polygon |>
    ggplot(mapping = aes(x = long, y = lat, group = group)) + 
    geom_polygon(aes(fill = as.factor(mean_urban_rural)), color = "black") + 
    labs(fill = "Average County Urban-Rural Code", caption = "Data Source: https://data.hrsa.gov/maps/mchb", title = "Average County Urban-Rural Code by State") +
    coord_map() + 
    theme_void() +
   scale_fill_viridis(discrete = TRUE) 
```
- This graph is a map of the continental US color-coded at the state-level by the mean county urban-rural code. The mean county urban-rural code ranges from 2 (most urban) to 6 (most rural). States on the east coast have the lowest urban-rural code, meaning that the average of state's counties' urban-rural codes is the most urban. States in the South and on the west coast are next-most urban. The states that are the most rural are North Dakota, South Dakota, Montana, and Wyoming. 





